Final draft

### 2021-07-19 final-draft.tex
commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Final draft},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{txfonts}
\usepackage{amsmath}
\usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{graphicx}
\usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
\usepackage{hyperref} %set the hyperlink to cite sections, equations...
\urlstyle{same}
\usepackage{comment}
\usepackage{enumitem}
\usepackage{ntheorem}
\usepackage{lipsum}
\usepackage{amssymb}
\usepackage{xcolor}
\theoremstyle{break}
\newtheorem{thm}{Theorem}% theorem counter resets every \subsection
\renewcommand{\thethm}{\arabic{thm}}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\theoremstyle{nonumberplain}
\newtheorem{proof*}{Proof.}
\newtheorem{definition}{Definition}
\renewcommand{\P}{\mathbb{P}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\setcounter{MaxMatrixCols}{10}
\setlength{\oddsidemargin}{.0in}
\setlength{\textwidth}{6 in}
\setlength{\topmargin}{-.1in}
\setlength{\textheight}{8.1in}
\renewcommand{\baselinestretch}{1.5}
\usepackage{amssymb}
\newcommand{\E}{\mathbb E}
\newcommand{\Q}{\mathbb Q}
\newcommand{\M}{\mathbb M}
\newcommand{\R}{\mathbb R}
\newcommand{\B}{\mathcal B}
\newcommand{\X}{\mathcal X}
\newcommand{\Pd}{\mathbb P}
\newcommand{\N}{\mathcal N}
\def\code#1{\texttt{#1}}
\newtheorem{algorithm}{Algorithm}[section]
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

We now show that State-Space models can be understood as Markov
processes that are partially observed. Let \(X_t\in\mathcal X\) and
\(Y_t\in\Y\) and let \(\{(X_t,T_t\}\) be a stochastic process. Consider
the initial distribution \(\mathbb{P}_0(dx_0)\) and the probability
kernels \(P_t(x_{t-1},dx_t)\) and \(F_t(x_t,d_t)\), where
\(t=1,2,\dots\).

\begin{definition}
The process $\{(X_t,T_t\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \mathbb{P}_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \mathbb{P}_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\mathbb{P}_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \mathbb{P}_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t)].
\end{align*}
\end{definition}

From the second inequality, it is apparent that \(\{(X_t,T_t\}\) is a
Markov process with initial distribution
\([\mathbb{P}_0(dx_0)F_0(x_0,y_0)]\) and transition kernel
\(P_t(x_{t-1},dx_t)F_t(x_t,dy_t).\) The third equality, instead, shows
that, \(\{X_t\}\) has a marginal distribution
\(\mathbb{P}_T(dx_{0:T})=\mathbb{P}_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),\)
meaning that \(\{X_t\}\) is marginally distributed as a Markov process
with initial distribution \(\mathbb{P}_T(dx_{0:T})\) and transition
kernel \(P_t(x_{t-1},dx_t).\)

@@

@@

Let us consider two probability spaces:
\((\mathcal X,\mathcal B(\mathcal X),\mathbb Q)\) and
\((\mathcal X,\mathcal B(\mathcal X),\mathbb M)\) and let the measurable
function \(w:\mathcal X\rightarrow \mathbb R\) be proportional to the
Radom-Nikodym derivative between \(\mathbb Q\) and \(\mathbb M\), i.e.,
\(w(x)\propto\frac{\mathbb Q(dx)}{\mathbb M(dx)}\).\footnote{From the Radom-Nikodym theorem a sufficient condition for $w$ to exist is that $\mathbb Q$ is absolutely continuous with respect to $\mathbb M$: $\mathbb Q\ll \mathbb M$.}\\
The following result represents the starting point for the idea of
importance sampling.

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Final draft},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{txfonts}
\usepackage{amsmath}
\usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{graphicx}
\usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
\usepackage{hyperref} %set the hyperlink to cite sections, equations...
\urlstyle{same}
\usepackage{comment}
\usepackage{enumitem}
\usepackage{ntheorem}
\usepackage{lipsum}
\usepackage{amssymb}
\usepackage{xcolor}
\theoremstyle{break}
\newtheorem{thm}{Theorem}% theorem counter resets every \subsection
\renewcommand{\thethm}{\arabic{thm}}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\theoremstyle{nonumberplain}
\newtheorem{proof*}{Proof.}
\newtheorem{definition}{Definition}
\renewcommand{\P}{\mathbb{P}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\setcounter{MaxMatrixCols}{10}
\setlength{\oddsidemargin}{.0in}
\setlength{\textwidth}{6 in}
\setlength{\topmargin}{-.1in}
\setlength{\textheight}{8.1in}
\renewcommand{\baselinestretch}{1.5}
\usepackage{amssymb}
\newcommand{\E}{\mathbb E}
\newcommand{\Q}{\mathbb Q}
\newcommand{\M}{\mathbb M}
\newcommand{\R}{\mathbb R}
\newcommand{\B}{\mathcal B}
\newcommand{\X}{\mathcal X}
\newcommand{\Y}{\mathcal Y}
\newcommand{\Pd}{\mathbb P}
\newcommand{\N}{\mathcal N}
\def\code#1{\texttt{#1}}
\newtheorem{algorithm}{Algorithm}[section]
\usepackage{float}
\floatplacement{figure}{}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

We now define \textit{State-Space models} as Markov processes that are
partially observed. Let \(X_t\in\mathcal X\) and \(Y_t\in\mathcal Y\)
and let \(\{(X_t,Y_t)\}\) be a stochastic process. To help intuition,
suppose that \(\{Y_t\}\) is observable, i.e., coincides with a sequence
of observations, while \(\{X_t\}\) is latent. Consider the initial
distribution \(\mathbb{P}_0(dx_0)\) and the probability kernels
\(P_t(x_{t-1},dx_t)\) and \(F_t(x_t,d_t)\), where \(t=1,2,\dots\).

\begin{definition}\label{state-space-def}
The process $\{(X_t,Y_t)\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \mathbb{P}_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \mathbb{P}_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\mathbb{P}_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \mathbb{P}_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t).
\end{align*}
\end{definition}

From the second inequality, it is apparent that \(\{(X_t,Y_t)\}\) is a
Markov process with initial distribution
\([\mathbb{P}_0(dx_0)F_0(x_0,y_0)]\) and transition kernel
\(P_t(x_{t-1},dx_t)F_t(x_t,dy_t).\) The third equality, instead, shows
that, \(\{X_t\}\) has a marginal distribution
\[\mathbb{P}_T(dx_{0:T})=\mathbb{P}_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),\]
meaning that \(\{X_t\}\) is marginally distributed as a Markov process
with initial distribution \(\mathbb{P}_T(dx_{0:T})\) and transition
kernel \(P_t(x_{t-1},dx_t).\)

Assume now that there exists a dominating measures \(\nu(dy)\) such that
\(F_t(x_t,dy_t)=f_t(y_t\mid x_t)\nu(dy_t)\) for all \(t.\) Then by the
last equality of definition \ref{state-space-def}, the joint
distribution of \((X_{0:T},Y_{0:T})\) is
\begin{equation}\label{state-space-densities}
   \mathbb{P}_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T})=\mathbb{P}_T(dx_{0:T})\prod\limits_{t=0}^T f_t(y_t\mid x_t)\prod\limits_{t=0}^T\vu(dy_t).
\end{equation} Integrating away \(X_{0:T},\) one gets the marginal
density \begin{equation}\label{state-space-marginal-y}
\begin{align*}
  \mathbb{P}_T(dy_{0:T}) &= \mathbb E_{\mathbb{P}_{t}}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big] \prod\limits_{s=0}^t\nu(dy_s) \\
                &= p_t(y_{0:t})\prod\limits_{s=0}^t\nu(dy_s),
\end{align*}                
\end{equation} where \(p_t(y_{0:t})\) is the likelihood function.
Dividing the left-hand side of equation \ref{state-space-densities} by
the left-hand side of equation \ref{state-space-marginal-y} and the
right-hand side of equation \ref{state-space-densities} by the
right-hand side of equation \ref{state-space-marginal-y}, one obtains
the filtering distribution \begin{equation}
\mathbb{P}_t(X_{t}\in dx_{t}\mid Y_{0:t}=y_{0:t})=\frac{1}{p_t(y_{0:t})}\big[\prod\limits_{s=0}^t f_s(y_s\mid X_s) \big]\mathbb{P}_t(dx_{0:t}).
\end{equation}

@@

\textit{Linear Gaussian models} are a class of popular state-space
models. A notable property of such models is that the filtering and
predictive distributions are analytically tractable, which makes such
models very convenient in applications.

Suppose that \(\mathcal X=\mathbb R^{d_x}\) and
\(\mathcal Y=\mathbb R^{d_y}.\) Let
\(U_t\thicksim\mathcal N(0,\Sigma_t)\) and
\(V_t\thicksim\mathcal N(0,R_t)\) be two independent white-noise errors
and assume the initial distribution
\(X_0\thicksim\mathcal N(0,\Sigma_0).\) We have a linear Gaussian model
if \begin{align}
X_t &= A_tX_{t-1}+U_t \\
Y_t &= B_tX_t + V_t.
\end{align} Note that \(X_t\mid X_{t-1}\) and \(Y_t\mid X_t\) are both
normally distributed.

\begin{lemma}\label{kalman-filter}
  Let $X\thicksim\mathcal N(m_0,Q_0)$ and $Y\mid X\thicksim\mathcal N(BX,R)$, where $R$ and $Q_0$ are positive semi-definite matrices. 
  Then, it holds that
  \begin{align}
    X\mid Y=y&\thicksim\mathcal N(m_1,Q_1),\\
    Q_1&= Q_0 (I_{d_x} - B'(B Q_0 B'+ R)^{-1} B Q_0) \\
    m_1&=[I_{d_x} - Q_0 B'(B Q_0 B'+ R)^{-1} B ]m_0 + Q_0 B'(B Q_0 B'+ R)^{-1}y.
  \end{align}
\end{lemma}
\begin{proof*}
   The lemma can be proved using Bayes' Theorem. Suppose $X\thicksim\mathcal N(m_0, Q_0).$ It is easy to see that
\begin{equation} \label{normal-distrib-rewrite}
        p(x)\propto \text{exp}\Big(-\frac{1}{2}X'AX + X'B\Big),
\end{equation}
with $A=Q_0^{-1}$ and $B=Am_0.$
Now, by Bayes' Theorem,

    \begin{align*}
        p(x\mid Y=y)  &\propto p(x) p(y\mid x) \\
                    &\propto \text{exp}\Big(-\frac{1}{2}X' Q_0^{-1}X + X' Q_0^{-1}m_0-\frac{1}{2}(y-BX)'R^{-1}(y-BX)) \\
                    &\propto \text{exp}(-\frac{1}{2}X' Q_0^{-1}X + X' Q_0^{-1}m_0-\frac{1}{2}X' B' R^{-1}BX+X' B' R^{-1}y)\\
                    &=\text{exp}(-\frac{1}{2}X'(Q_0^{-1}+B' R^{-1}B)X+
                    X'(Q_0^{-1}m_0+B' R^{-1}y)\Big).
    \end{align*}

Hence, it follows that 
\begin{equation*}
    Q_1=Var(X\mid Y=y)= (Q_0^{-1}+ B' R^{-1} B)^{-1}=Q_0-Q_0 B'(B Q_0 B'+R)^{-1}B Q_0
\end{equation*}
and 
    \begin{align*}
        m_1 &=\mathbb E(X\mid Y=y) = Q_1^{-1}(Q_0^{-1}m_0+B' R^{-1}y)\\
            &= (Q_0-Q_0B'(B Q_0B'+R)^{-1}B Q_0)(Q_0^{-1}m_0+B' R^{-1}y)\\
            &= m_0-Q_0B'(B Q_0B'+R)^{-1}B m_0 + Q_0 B' R^{-1}y - Q_0B'(B Q_0B'+R)^{-1}B Q_0 B' R^{-1}y \\
            &=[I_{d_x} - Q_0 B'(B Q_0 B'+ R)^{-1} B ]m_0 + Q_0 B'(B Q_0 B'+ R)^{-1}y.
    \end{align*}
\end{proof*}

Assume that
\(\mathbb{P}_{t-1}(dx_{t-1}\mid Y_{0:t-1}=y_{0:t-1})=\mathcal N(m_{t-1},Q_{t-1}).\)
Then, using the transition equation of the latent variable, by the
linearity of the Gaussian distribution, we have that the predictive
distribution is \begin{equation}
\mathbb{P}_{t-1}(dx_{t}\mid Y_{0:t-1}=y_{0:t-1})=\mathcal N(A_tm_{t-1},A_tQ_{t-1}A_t'+\Sigma_t).
\end{equation} Letting \(E_T=A_tQ_{t-1}A_t'+\Sigma_t,\) and applying
lemma \ref{kalman-filter}, it follows that the filtering distribution is
\begin{align*}
  \mathbb{P}_{t}(X_t\in dx_{t}\mid Y_{0:t-1}&=y_{0:t-1})=\mathcal N(m_t,Q_t), \text{where} \\ Q_t&=E_t[I_{d_x} - B_t'(B_tE_tB_t'+R_t)^{-1} B_tE_t] \\
  m_t &= [I_{d_x} - E_tB_t'(B_tE_tB_t'+R_t)^{-1} B_t]A_tm_{t-1}+E_tB_t'(B_tE_tB_t'+R_t)^{-1}y_t.
\end{align*}

\section{Implementation}

In this section we present a practical illustration of the algorithms
discussed. For the sake of simplicity we use a random walk plus noise
model, i.e.~the most basic form of a linear Gaussian state-space
model.\\
\begin{align}
y_{t}|x_{t} & \sim N(x_{t},\sigma^{2}) \\
x_{t}|x_{t-1} & \sim N(x_{t-1},\tau^{2}) \\
x_{0} & \sim N(m_{0},C_{0})
\end{align} As already mentioned before, in this case the filtering
distribution can be computed in closed form solutions using the Kalman
filter. However, this toy example will be used also to illustrate more
involved filtering strategies described in this work. We believe indeed
that it represents a useful starting point to understand the logic of
the algorithms which may be eventually replicated when dealing with more
complex models.\\
The filtering strategy is applied to 50 simulated data. Figure XX shows
the simulated true states sequence assuming as data generating process
the Equation (2) with \(\tau^{2}=1\) and simulated observed sequence
process form Equation (1) with \(\sigma^{2}=1\).

{\centering \includegraphics{final-draft_files/figure-latex/unnamed-chunk-5-1}

\caption{Simulated random walk plus noise model}\label{fig:unnamed-chunk-5}
\end{figure}

The Kalman Filter for this model is implemented as described below.

\begin{algorithm} Kalman Filter for Random Walk plus Noise Model
\begin{itemize}
\item Initialize $\theta_{0} \sim N(m_{0},C_{0})$
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute the one-step-ahead state predictive distribution at time $t-1$, notice that if $t=1$ no data have been observed yet and therefore $x_{1}|x_{0}\sim N(a_{1},R_{1})$ otherwise
\begin{align*}
x_{t}|y_{1:t-1} & \sim N(a_{t},R_{t})\\
a_{t} & = m_{t-1}\\
R_{t} & = C_{t-1}+\tau^2
\end{align*}
\item Compute the filtering distribution at time $t$ as $p(x_{t}|y_{1:t}) \propto p(x_{t}|y_{1:t-1})p(y_{t}|x_{t})$, i.e. the product of the one-step-ahead state predictive distribution and the likelihood
\begin{align*}
x_{t}|y_{1:t} & \sim N(m_{t},C_{t}) \\
m_{t} & = \big(1-\frac{R_{t}}{R_{t}+\sigma^2}\big)a_{t}+\frac{R_{t}}{R_{t}+\sigma^2}y_{t} \\
C_{t} & = \frac{R_{t}}{R_{t}+\sigma^2}\sigma^2
\end{align*}
\end{enumerate}
\end{itemize}
\end{algorithm}

Our \texttt{DLM} function implement in R replicate this steps.\\

\hrule
\hrule
\texttt{DLM(data,sig2,tau2,m0,C0)}
\hrule

\textbf{Arguments}

\texttt{data} ~~the observed process. It has to be a vector or a
univariate time series.\\
\texttt{sig2} ~~the variance \(\sigma^{2}\) in Equation (1)\\
\texttt{tau2} ~~the variance \(\tau^{2}\) in Equation (2)\\
\texttt{m0} ~~central value of the normal prior state distribution\\
\texttt{C0} ~~variance of the normal prior state distribution

\hrule
\hrule

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DLM}\OtherTok{\textless{}{-}}\ControlFlowTok{function}\NormalTok{(data,sig2,tau2,m0,C0)\{}
\NormalTok{  n  }\OtherTok{=} \FunctionTok{length}\NormalTok{(data)}
\NormalTok{  m  }\OtherTok{=} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,n)}
\NormalTok{  C  }\OtherTok{=} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,n)}
  \ControlFlowTok{for}\NormalTok{ (t }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{n)\{}
    \ControlFlowTok{if}\NormalTok{ (t}\SpecialCharTok{==}\DecValTok{1}\NormalTok{)\{}
\NormalTok{      a }\OtherTok{=}\NormalTok{ m0}
\NormalTok{      R }\OtherTok{=}\NormalTok{ C0 }\SpecialCharTok{+}\NormalTok{ tau2}
\NormalTok{    \}}\ControlFlowTok{else}\NormalTok{\{}
\NormalTok{      a }\OtherTok{=}\NormalTok{ m[t}\DecValTok{{-}1}\NormalTok{]}
\NormalTok{      R }\OtherTok{=}\NormalTok{ C[t}\DecValTok{{-}1}\NormalTok{] }\SpecialCharTok{+}\NormalTok{ tau2}
\NormalTok{    \}}
\NormalTok{    A }\OtherTok{=}\NormalTok{ R}\SpecialCharTok{/}\NormalTok{(R}\SpecialCharTok{+}\NormalTok{sig2)}
\NormalTok{    m[t] }\OtherTok{=}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{A)}\SpecialCharTok{*}\NormalTok{a }\SpecialCharTok{+}\NormalTok{ A}\SpecialCharTok{*}\NormalTok{y[t]}
\NormalTok{    C[t] }\OtherTok{=}\NormalTok{ A}\SpecialCharTok{*}\NormalTok{sig2}
\NormalTok{  \}}
  \FunctionTok{return}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{m=}\NormalTok{m,}\AttributeTok{C=}\NormalTok{C))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

In Figure XX below filtered states estimated using Kalman Filter with
\(x_{0} \sim N(0,100)\) and \(\sigma^{2}=\tau^{2}=1\) are compared to
the true states values. Notice how closely the filtered states follow
the observations and the goodness of the approximation of the true
states. 95 percent credible intervals are computed as
\[[E(\theta_{t}|y_{1:t})-z_{1-\alpha/2}\sqrt{V(\theta_{t}|y_{1:t})},E(\theta_{t}|y_{1:t})+z_{1-\alpha/2}\sqrt{V(\theta_{t}|y_{1:t})}]\]

{\centering \includegraphics{final-draft_files/figure-latex/unnamed-chunk-7-1}

\caption{Kalman Filtered States with credible interval (in red)}\label{fig:unnamed-chunk-7}
\end{figure}

The discussion of Kalman Filter will continue in Section XX where we
compare it to the Particle Filter.

\textbf{Implementation}

In this section we present a practical illustration of the algorithms
discussed. For the sake of simplicity we use a random walk plus noise
model, i.e.~the most basic form of a linear Gaussian state-space model.
\begin{align}
y_{t}|x_{t} & \sim N(x_{t},\sigma^{2}) \\
x_{t}|x_{t-1} & \sim N(x_{t-1},\tau^{2}) \\
x_{0} & \sim N(m_{0},C_{0})
\end{align} As already mentioned before, in this case the filtering
distribution can be computed in closed form solutions using the Kalman
filter. However, this toy example will be used also to illustrate more
involved filtering strategies described in this work. We believe indeed
that it represents a useful starting point to understand the logic of
the algorithms which may be eventually replicated when dealing with more
complex models.\\
The filtering strategy is applied to 50 simulated data. Figure XX shows
the simulated true states sequence assuming as data generating process
the Equation (2) with \(\tau^{2}=1\) and simulated observed sequence
process form Equation (1) with \(\sigma^{2}=1\).

{\centering \includegraphics{final-draft_files/figure-latex/unnamed-chunk-8-1}

\caption{Simulated random walk plus noise model}\label{fig:unnamed-chunk-8}
\end{figure}

The discussion of Kalman Filter will continue in Section XX where we
compare it to the Particle Filter.

@@

Let us consider two probability spaces:
\((\mathcal X,\mathcal B(\mathcal X),\mathbb Q)\) and
\((\mathcal X,\mathcal B(\mathcal X),\mathbb M)\) and let the measurable
function \(w:\mathcal X\rightarrow \mathbb R\) be proportional to the
Radom-Nikodym derivative between \(\mathbb Q\) and \(\mathbb M\), i.e.,
\(w(x)\propto\frac{\mathbb Q(dx)}{\mathbb M(dx)}\).\\
The following result represents the starting point for the idea of
importance sampling.

@@

\section{Implementation}

Consider again the random walk plus noise model of section XX,the
challenge here is to estimate the filtered states using a Sequential
Importance Sampling algorithm. The main idea of the SIS applied to this
univariate linear gaussian model is described in the following
algorithm. We indicate with \(n\) the sample size of time observations
and with \(N\) the generated sample size for each step of the Sequenial
Monte Carlo.

\begin{algorithm} SIS filter for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}

Our \texttt{SISfun} function implemented in R replicate this steps.\\
